DeepFovea++: Reconstruction and Super-Resolution for Natural Foveated Rendered Videos

### 2020-07-22 DeepFovea++ paper/paper.tex
Update paper

\begin{abstract} \label{abstract}
Image super-resolution is a well-known problem in the computer vision community. Recent papers extended the problem of super-resolution to videos and showed amazing results. On the other hand deep learning based fovea sampled image reconstruction has drawn some popularity since the DeepFovea publication of Facebook AI. Even though DeepFovea showed outstanding results, the proposed reconstruction network was only able to reconstruct relatively low-resolution images of $128\times128$ pixels. We revisit the proposed DeepFovea architecture to perform fovea sampled video reconstruction and super-resolution ($192 \times 256$ $\to$ $768 \times 1024$) at once. Our proposed architecture, DeepFovea++, first reconstructs a given video sequence by a recurrent U-Net architecture, and afterwards, the desired super-resolution is learned by deformable convolutions. We tested our DeepFovea++ architecture on the challenging REDS dataset. The code is available at {\url{https://github.com/ChristophReich1996/DeepFoveaPP_for_Video_Reconstruction_and_Super_Resolution}}.

@@

\section{Introduction} \label{sec:introduction}
A full immersion with virtual reality requires a very high image resolution, a low latency, and a high frame refresh rate \cite{deepfovea}. Kaplanyan \etal \cite{deepfovea} from Facebook Reality Labs approached this problem by making use of the fact that human perception has different levels of visual acuity across the retina. The closer regions in the field of view are to the point of maximum focus - the fovea centralis - the more details from the incoming visual image are perceived. This makes many of the high-level details in image regions that are further away from the fovea point unnecessary, as they are not recognized by the human eye. Instead, the human brain infers this missing visual information based on experience it has made in the past.\\
The authors of DeepFovea \cite{deepfovea} adopted this concept by dropping some of the peripheral pixels for the sake of reduced computational effort. They proposed a deep video reconstruction model that takes such foveated video frames and generates a plausible hypothesis for the missing peripheral pixels based on the remaining ones.\\
In this foveated rendering setup, recorded frames can be masked so that some of the pixels are dropped w.r.t. a probability that is higher the further away the respective pixel is from center of focus. This new masked frame requires less space. They were able to achieve a compression rate of up to 14x without a significant loss in perceived video quality.\\
However, as far as we know, the authors of the DeepFovea paper were working on relatively small images with a resolution of $128\times 128$ pixels. This motivated us to go a step further and add a video super resolution extension to their basic DeepFovea architecture. We trained it in an end-to-end fashion and were able to generate upscaled and reconstructed output frames of a four times higher resolution.\\

@@

Where $\hat{\Tensor{I}}$ is the reconstructed super-resolution image and $D$ the 3d discriminator or the 3d FFT discriminator. The discriminator networks are trained on the corresponding counter part to the non-saturating generator loss, which is defined as $-\Mean{\log\left(\operatorname{D}\left(\Tensor{I}\right)\right)}-\Mean{\log\left(1-\operatorname{
D}\left(\hat{\Tensor{I}}\right)\right)}$. Where $\Tensor{I}$ is the super-resolution ground truth label. \cite{gan}\\
Besides the supervised and adversarial losses, we also utilized the calibrated perceptual loss (LPIPS) \cite{perceptual} to achieve more natural-looking reconstructed images.

\begin{equation}
    \mathcal{L}_{\text{LPIPS}} = \frac{1}{5}\sum_{i=1}^{5}\frac{1}{c_{i}\,h_{i}\,w_{i}}\norm{\operatorname{VGG}_{i,2}\left(\hat{\Tensor{I}}\right) - \operatorname{VGG}_{i,2}\left(\Tensor{I}\right)}_{1}
\end{equation}

@@

\subsection{REDS Dataset} \label{subsec:redsdataset}
In our experiments, we used the REDS dataset \cite{REDS}, which has been compiled in 2019 mainly for the tasks of video deblurring and video super-resolution. It totals $300$ sequences of $100$ high-quality RGB video frames each, with a resolution of $720\times 1280$. These sequences are split into $240$ videos for training, $30$ for validation, and $30$ for testing. However, only the training and test set are publicly available (as of July, 2020). In order to get input-output example pairs for supervised training, we followed the authors of the dataset by downscaling the target high-resolution images by a factor of $4$ to obtain the unmasked low-resolution input frames. However, we used bi-linear instead of bi-cubic interpolation. \cite{REDS}

@@

$p_{\Vector{x}}$ represents the probability that the pixel of the image $\Matrix{I}\in\Set{R}^{3\times\text{height}\times\text{width}}$ at the given position $\Vector{x}=[x_{0}, x_{1}]\in\Set{R}^2$ is not masked out. We mask out the image over all three RGB channels. The vector $\Vector{x}_{f}=[x_{0,f}, x_{1,f}]\in\Set{R}^2$ represent the focus point of the fovea mask. This focus point is randomly generated by 2d uniform distribution which is defined as $\mathcal{U}\left(\left[50, \text{height} - 50\right]; \left[50, \text{width} - 50\right]\right)$. To generate the final fovea mask we sample for each pixel from a uniform distribution $\mathcal{U}\left(0; 1\right)$ and threshold the resulting value with the corresponding $p_{\Vector{x}}$ value to generate a binary mask. This binary mask is then applied to the input image.\\
The preprocessing in our asynchronous data loader contains five steps. In the first step, a batch of high-resolution images ($720 \times 1280$) is loaded to construct a sequence of six frames. Then the images are downscaled by bi-linear interpolation to reach the resolution of $180 \times 320$ pixels. Next, both the inputs and the high-resolution labels are normalized to achieve pixel values between zero and one. Afterward, the fovea mask is generated and applied to the input image. In the final step, the fovea sampled images and the high-resolution label images are cropped and padded to a resolution of $192 \times 256$ and $768 \times 1024$, respectively. This is necessary to match the required input shape of the reconstruction network.\\
We set the weight factors of the loss function \ref{eq:loss} to balance out the magnitudes between the different loss functions. The following weights were used: $w_{\text{sv}}=5$, $w_{\text{adv}}=0.01$, $w_{\text{adv ff}}=0.01$, $w_{\text{flow}}=2$ and $w_{\text{LPIPS}}=1$.
We were able to fit one sequence consisting of six RGB frames on one Tesla V100 (16GB), at training time. We trained our whole framework for 15 epochs on two GPUs with a batch size of two. The training process took us about two days.

\subsection{Results} \label{subsec:results}
For analyzing our framework, results we calculated multiple metrics. First, we compute the common $\operatorname{L1}$ metric and the Mean-Squared-Error (MSE, $\operatorname{L2}$). In case of a reconstructed image $\hat{\Tensor{I}}\in\Set{R}^{c, h, w}$ and the corresponding label $\Tensor{I}\in\Set{R}^{c, h, w}$, the $\operatorname{L1}$ and $\operatorname{L2}$ loss is defined as

\begin{align}
    \operatorname{L1} &= \frac{1}{c\, h\, w}\norm{\hat{\Tensor{I}} - \Tensor{I}}_{1}\\
    \operatorname{L2} &= \frac{1}{c\, h\, w}\norm{\hat{\Tensor{I}} - \Tensor{I}}_{2}.
\end{align}

>>>>

	\begin{abstract} \label{abstract}
		Image super-resolution is a well-known problem in the computer vision community. Recent papers extended the problem of super-resolution to videos and showed amazing results. On the other hand deep learning based fovea sampled image reconstruction has drawn some popularity since the DeepFovea publication of Facebook AI. Even though DeepFovea showed outstanding results, the proposed reconstruction network was only able to reconstruct relatively low-resolution frames of $128\times128$ pixels. We revisit the proposed DeepFovea architecture to perform fovea sampled video reconstruction and super-resolution ($192 \times 256$ $\to$ $768 \times 1024$) at once. Our proposed architecture, DeepFovea++, first reconstructs a given video sequence by a recurrent U-Net architecture, and afterwards, the desired super-resolution is learned by deformable convolutions. We tested our DeepFovea++ architecture on the challenging REDS dataset. The code is available at {\url{https://github.com/ChristophReich1996/DeepFoveaPP_for_Video_Reconstruction_and_Super_Resolution}}.

@@

	\section{Introduction} \label{sec:introduction}
	A full immersion with virtual reality requires a very high image resolution, a low latency, and a high frame refresh rate \cite{deepfovea}. Kaplanyan \etal \cite{deepfovea} from Facebook Reality Labs approached this problem by making use of the fact that human perception has different levels of visual acuity across the retina. The closer regions in the field of view are to the point of maximum focus - the fovea centralis - the more details from the incoming visual image are perceived. This makes many of the high-level details in image regions that are further away from the fovea point unnecessary, as they are not recognized by the human eye. Instead, the human brain infers this missing visual information based on experience it has made in the past.\\
	The authors of DeepFovea \cite{deepfovea} adopted this concept by dropping some of the peripheral pixels for the sake of reduced computational effort. They proposed a deep video reconstruction model that takes such foveated video frames and generates a plausible hypothesis for the missing peripheral pixels based on the remaining ones.\\
	In this foveated rendering setup, recorded frames can be masked so that some of the pixels are dropped w.r.t. a probability that is higher the further away the respective pixel is from center of focus. This new masked frame requires less space. They were able to achieve a compression rate of up to 14x without a significant loss in perceived video quality.\\
	However, as far as we know, the authors of the DeepFovea paper were working on relatively small video frames with a resolution of $128\times 128$ pixels. This motivated us to go a step further and add a video super resolution extension to their basic DeepFovea architecture. We trained it in an end-to-end fashion and were able to generate upscaled and reconstructed output frames of a four times higher resolution.\\

@@

	Where $\hat{\Tensor{I}}$ is the reconstructed super-resolution video sequence and $D$ the 3d discriminator or the 3d FFT discriminator. The discriminator networks are trained on the corresponding counter part to the non-saturating generator loss, which is defined as $-\Mean{\log\left(\operatorname{D}\left(\Tensor{I}\right)\right)}-\Mean{\log\left(1-\operatorname{
			D}\left(\hat{\Tensor{I}}\right)\right)}$. Where $\Tensor{I}$ is the super-resolution ground truth label. \cite{gan}\\
	Besides the supervised and adversarial losses, we also utilized the calibrated perceptual loss (LPIPS) \cite{perceptual} to achieve more natural-looking reconstructed frames.

	\begin{equation}
	\mathcal{L}_{\text{LPIPS}} = \frac{1}{5}\sum_{i=1}^{5}\frac{1}{b\,c_{i}\,h_{i}\,w_{i}}\norm{\operatorname{VGG}_{i,2}\left(\hat{\Tensor{I}}\right) - \operatorname{VGG}_{i,2}\left(\Tensor{I}\right)}_{1}
	\end{equation}

@@

	\subsection{REDS Dataset} \label{subsec:redsdataset}
	In our experiments, we used the REDS dataset \cite{REDS}, which has been compiled in 2019 mainly for the tasks of video deblurring and video super-resolution. It totals $300$ sequences of $100$ high-quality RGB video frames each, with a resolution of $720\times 1280$. These sequences are split into $240$ videos for training, $30$ for validation, and $30$ for testing. However, only the training and test set are publicly available (as of July, 2020). In order to get input-output example pairs for supervised training, we followed the authors of the dataset by downscaling the target high-resolution video frames by a factor of $4$ to obtain the unmasked low-resolution input frames. However, we used bi-linear instead of bi-cubic interpolation. \cite{REDS}

@@

	$p_{\Vector{x}}$ represents the probability that the pixel of the image $\Matrix{I}\in\Set{R}^{3\times\text{height}\times\text{width}}$ at the given position $\Vector{x}=[x_{0}, x_{1}]\in\Set{R}^2$ is not masked out. We mask out the image over all three RGB channels. The vector $\Vector{x}_{f}=[x_{0,f}, x_{1,f}]\in\Set{R}^2$ represent the focus point of the fovea mask. This focus point is randomly generated by 2d uniform distribution which is defined as $\mathcal{U}\left(\left[50, \text{height} - 50\right]; \left[50, \text{width} - 50\right]\right)$. To generate the final fovea mask we sample for each pixel from a uniform distribution $\mathcal{U}\left(0; 1\right)$ and threshold the resulting value with the corresponding $p_{\Vector{x}}$ value to generate a binary mask. This binary mask is then applied to the input image.\\
	The preprocessing in our asynchronous data loader contains five steps. In the first step, a batch of high-resolution frames ($720 \times 1280$) is loaded to construct a sequence of six frames. Then the frames are downscaled by bilinear interpolation to reach the resolution of $180 \times 320$ pixels. Next, both the inputs and the high-resolution labels are normalized to achieve pixel values between zero and one. Afterward, the fovea mask is generated and applied to the input image. In the final step, the fovea sampled images and the high-resolution label images are cropped and padded to a resolution of $192 \times 256$ and $768 \times 1024$, respectively. This is necessary to match the required input shape of the reconstruction network.\\
	We set the weight factors of the loss function \ref{eq:loss} to balance out the magnitudes between the different loss functions. The following weights were used: $w_{\text{sv}}=5$, $w_{\text{adv}}=0.01$, $w_{\text{adv ff}}=0.01$, $w_{\text{flow}}=2$ and $w_{\text{LPIPS}}=1$.
	We were able to fit one sequence consisting of six RGB frames on one Tesla V100 (16GB), at training time. We trained our whole framework for 15 epochs on two GPUs with a batch size of two. The training process took us about two days.

	\subsection{Results} \label{subsec:results}
	For analyzing our framework, results we calculated multiple metrics. First, we compute the common $\operatorname{L1}$ metric and the Mean-Squared-Error (MSE, $\operatorname{L2}$). In case of a reconstructed image sequence $\hat{\Tensor{I}}\in\Set{R}^{b\,c, h, w}$ and the corresponding label $\Tensor{I}\in\Set{R}^{b\,c, h, w}$, the $\operatorname{L1}$ and the $\operatorname{L2}$ loss is defined as

	\begin{align}
	\operatorname{L1} &= \frac{1}{b\,c\, h\, w}\norm{\hat{\Tensor{I}} - \Tensor{I}}_{1}\\
	\operatorname{L2} &= \frac{1}{b\,c\, h\, w}\norm{\hat{\Tensor{I}} - \Tensor{I}}_{2}.
	\end{align}
